Sparse Function-space Representation \\ of Neural Networks

### 2023-05-23 paper/main.tex
App. D.3

\subsection{Uncertainty quantification on UCI data sets}
\label{app:uci}
%
In \cref{sec:uci} in the main paper, we showed experiments on 8 UCI data sets \cite{UCI} matching those presented in \cite{immer2021improving}. For direct comparison, we replicate the benchmark as in \cite{immer2021improving}: for each data set, we run the neural network training and the \our prediction pipeline on 10 values of prior precision $\delta$ and 10 seeds matching those used in \cite{immer2021improving}. After completing these 100 runs, we select the $\delta$ with the lowest mean validation set NLPD and report the test set performance statistics in our result tables. The range of $\delta$ used for the hyperparameter search varies depending on data set and the number of inducing points. Below we list the selected $\delta$ ranges:
\begin{itemize}
\item {\sc digits}: The ten $\delta$ values are logspaced between $10^{-1}$ and $10^2$, matching~\cite{immer2021improving}.
\item {\sc cancer}: For $M > 128$, the ten $\delta$ values are logspaced between $10^{-1}$ and $10^2$. With too low values of $\delta$, we observed numerical errors in the Cholesky decomposition required for constructing the \our dual.
\item {\sc satellite}: For $M < 512$, the tend $\delta$ values are logspaced between $10^{-1}$ and $10^2$, matching~\cite{immer2021improving}.
\item {\sc satellite}: For $M \geq 512$, the ten $\delta$ values are logspaced between $10^{1}$ and $10^2$.  With too low values of $\delta$, we observed numerical errors in the Cholesky decomposition required for constructing the \our dual.
\item Othewise: The ten $\delta$ values are logspaced between $10^{-2}$ and $10^2$, matching~\cite{immer2021improving}.
\end{itemize}

@@

\subsection{Uncertainty benchmarks for image data}
\label{app:image}
% Each experiment with the image data sets consisted of the following steps.
% First, we trained a neural network (MLP for MNIST, CNN for FMNIST) using a prior precision $\delta$.
% After training, we performed inference by calculating the dual parameters.
% Finally, we evaluated the predictive performance of \our on the validation set.
% Similarly to the supervised UCI benchmark tasks, a single image experiment consists of the following steps: first, we train a neural network (a MLP for MNIST, a CNN for FMNIST) in the classification task, using a prior precision $\delta$. After training, we prepare the \our method for predictions by computing the dual parameters, and then evaluate the validation set predictive performance of \our. While in principle our method's predictions consist of the GP mean and variance, we include results for using the neural network prediction as a mean instead. For comparison, we gather the GP subset method results for both the GP and neural network means. This experiment set-up is then replicated for a range of prior precision values and learning rates, as we perform a grid search over these hyperparameters by optimizing for the validation set NLPD. We run the experiment on $5$ seeds and pick the hyperparameter by the best mean performance over the seeds. The results reported in \cref{tbl:imagesuper} used a learning rate of $10^{-4}$, and a prior precision of $\delta = 0.0013$ for FMNIST and $\delta = 0.0001$ for MNIST \todo{double check}.

@@

\subsection{\our for continual learning}
\label{app:cl-exp}
\textbf{CL Data sets}
We conducted our experiments on three common CL data sets \cite{de2021continual, pan2020continual, rudner2022continual}: 
\begin{itemize}
	\item[-] Split-MNIST (S-MNIST), that consists in five binary classification tasks each on pair on MNIST classes (\ie, \digit{0}~and~\digit{1}, \digit{2}~and~\digit{3}, \dots, \digit{8}~and~\digit{9});
	\item[-] Split-FashionMNIST (S-FMNIST), composed of five binary tasks as S-MNIST, but using instead Fashion-MNIST;
	\item[-] Permuted-MNIST (P-MNIST), that consists of ten tasks, where each one is a ten-way classification task on the entire MNIST data set, where the images' pixels are randomly permuted.
\end{itemize}

Following \cite{pan2020continual, rudner2022continual}, we used a two-layer MLP with 256 hidden units and ReLU activations for S-MNIST and S-FMNIST and a two-layer MLP with 100 hidden units and ReLU activations for P-MNIST, to ensure consistency with the other methods. In our experiments we only consider the single-head setting (SH). This means that for S-MNIST and S-FMNIST, we are dealing with a Class-Incremental Learning (CIL) scenario, where for each task a pair of dijoint classes are presented to the model. The P-MNIST experiments falls instead into the Domain-Incremental Learning (DIL) setting, where the model deals with a data-distribution shift, by progressively observing tasks with different permutations of the image pixels that need to be classified into ten different classes.

In our experiments, we compared the performance of \our-based regularizer against two families of methods: weight-space regularization methods and function-space regularization ones.
For the weight-space methods, we considered Online-EWC~\citep{schwarz2018progress}, SI~\citep{zenke2017continual}, and VCL~\citep{nguyen2018variational}.  % In \cref{tbl:cl_table_1}, we considered VCL with random coresets, which uses coreset points t 
For the function-space methods, we considered DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual}.

For our method, DER, Online-EWC, and SI, we relied on the Mammoth framework proposed by \citep{buzzega2020dark}. On the other hand, we used the available open-source FROMP codebase to conduct the experiments with FROMP, and the S-FSVI codebase to run the experiments on S-FSVI and VCL. We selected the hyperparameters to use with all these methods trying to adapt the ones pointed out in the their original papers. For this reason, we relied on SGD for DER, Online-EWC, and SI, which requires higher learning rates, compared to the others methods for which we used Adam, \ie, our \our regularizer, FROMP, S-FSVI and VCL.

For the methods relying on a stored subset of training data for each task, we make sure that the same amount of data points store are the same for each method. For DER, which does not select the rehearsal training samples at task boundaries, we cannot apply directly the definition of points per task ($M$), and then we set the dimension of the reservoir buffer to be $T \cdot M$, \ie, total number of tasks times the points per task.

The results reported in \cref{tbl:cl_table_1}, are obtained using the following hyperparameters:

%% Hyperparms used in the different experiments:
\textbf{S-MNIST (40 pts./task):}
\begin{itemize}
	\item[-] Online-EWC: SGD, learning rate $0.03$, batch size  10, number of epochs 1, $\lambda=90$, $\gamma=1.0$.
	\item[-] SI: SGD, learning rate 0.1, batch size 10, number of epochs 1, $c=1.0$, $\xi=0.9$.
	\item[-] VCL (with coresets): Adam, learning rate 0.001, batch size 256, number of epochs 100, selection method for coresets random choice.
	\item[-] DER: SGD with learning rate 0.03, batch size 10, buffer batch size 10, number epochs 1, $\alpha=0.3$, reservoir buffer size 200.
	\item[-] FROMP: Adam, learning rate $1 \cdot 10^{-4}$, batch size 32, number of epochs 10, $\tau = 10$, random selection for memorable points.
	\item[-] S-FSVI: Adam, learning rate $5 \cdot 10^{-4}$, batch size 128, number of epochs 10.
	\item[-] SFR: Adam, learning rate $3 \cdot 10^{-4}$, batch size 32, $\tau=1$, $\delta=0.0001$,  number of epochs 1.
\end{itemize}

\textbf{S-MNIST (200 pts./task):}
\begin{itemize}
	\item[-] Online-EWC: SGD, learning rate $0.03$, batch size  10, number of epochs 1, $\lambda=90$, $\gamma=1.0$.
	\item[-] SI: SGD, learning rate 0.1, batch size 10, number of epochs 1, $c=1.0$, $\xi=0.9$.
	\item[-] VCL (with coresets): Adam, learning rate 0.001, batch size 256, number of epochs 100, selection method for coresets random choice.
	\item[-] DER: SGD, learning rate 0.03, batch size 10, buffer batch size 10, number epochs 1, $\alpha$ 0.2, reservoir buffer size 1000.
	\item[-] FROMP: Adam, learning rate $1 \cdot 10^{-4}$, batch size 32, number of epochs 10, $\tau = 10$, random selection for memorable points.
	\item[-] S-FSVI: Adam, learning rate $5 \cdot 10^{-4}$, batch size 128, number of epochs 80.
	\item[-] SFR: Adam, learning rate $1 \cdot 10^{-4}$, batch size 32, $\tau=0.5$, $\delta=1 \cdot 10^{-5}$, number of epochs=5.
\end{itemize}

%% Placeholders for now
\textbf{S-FMNIST (200 pts./task):}
\begin{itemize}
	\item[-] Online-EWC: SGD, learning rate $0.03$, batch size  10, number of epochs 5, $\lambda=90$, $\gamma=1.0$.
	\item[-] SI: SGD, learning rate 0.1, batch size 10, number of epochs 5, $c=1.0$, $\xi=0.9$.
	\item[-] VCL (with coresets): Adam, learning rate $1 \cdot 10^{-3}$, batch size 256, number of epochs 100, selection method for coresets random choice.
	\item[-] DER: SGD, learning rate 0.03, batch size 10, buffer batch size 10, number epochs 5, $\alpha =  0.2$, reservoir buffer size 200.
	\item[-] FROMP: Adam, learning rate $1 \cdot 10^{-4}$, batch size 32, number of epochs 10, $\tau = 10$, random selection for memorable points.
	\item[-] S-FSVI: Adam, learning rate $5 \cdot 10^{-4}$, batch size 128, number of epochs 60.
	\item[-] SFR: Adam, learning rate $3 \cdot 10^{-4}$, batch size 32, $\tau=1.0$, $\delta=0.0001$, number of epochs=5.
\end{itemize}

\textbf{P-MNIST (200 pts./task):}
\begin{itemize}
	\item[-] Online-EWC: SGD, learning rate $0.1$, batch size 128, number of epochs 10, $\lambda=0.7$, $\gamma=1.0$.
	\item[-] SI: SGD, learning rate 0.1, batch size 138, number of epochs 10, $c=0.5$, $\xi=1.0$. 
	\item[-] VCL (with coresets): Adam, learning rate 0.001, batch size 256, number of epochs 100, selection method for coresets random choice.
	\item[-] DER:  SGD, learning rate 0.01, batch size 128, buffer batch size 128, number epochs 10, $\alpha=0.3$, reservoir buffer size 2000.
	\item[-] FROMP: Adam, learning rate $1 \cdot 10^{-3}$, batch size 128, number of epochs 10, $\tau = 0.5$, lambda descend selection for memorable points.
	\item[-] S-FSVI: Adam, learning rate $5 \cdot 10^{-4}$, batch size 128, number of epochs 10.
	\item[-] SFR: Adam, learning rate $3 \cdot 10^{-4}$, batch size 64, $\tau=1.0$, $\delta=0.0001$. 
\end{itemize}

In addition the ablation experiments for SFR, we used the same hyperparameters reported for the full approach.

>>>>

\subsection{Capturing uncertainty in UCI tasks under supervised learning}
\label{app:uci}
%
In \cref{sec:uci} in the main paper, we showed experiments on 8 UCI data sets \cite{UCI} matching those presented in \cite{immer2021improving}. For direct comparison, we replicate the benchmark as in \cite{immer2021improving}: for each data set, we run the neural network training and the \our prediction pipeline on 10 values of prior precision $\delta$ and 10 seeds matching those used in \cite{immer2021improving}. After completing these 100 runs, we select the $\delta$ with the lowest mean validation set NLPD and report the test set performance statistics in our result tables. The range of $\delta$ used for the hyperparameter search varies depending on data set and the number of inducing points. Below we list the selected $\delta$ ranges:
\begin{itemize}
\item {\sc digits}: The ten $\delta$ values are logspaced between $10^{-1}$ and $10^2$, matching~\cite{immer2021improving}.
\item {\sc cancer}: For $M > 128$, the ten $\delta$ values are logspaced between $10^{-1}$ and $10^2$. With too low values of $\delta$, we observed numerical errors in the Cholesky decomposition required for constructing the \our dual.
\item {\sc satellite}: For $M < 512$, the tend $\delta$ values are logspaced between $10^{-1}$ and $10^2$, matching~\cite{immer2021improving}.
\item {\sc satellite}: For $M \geq 512$, the ten $\delta$ values are logspaced between $10^{1}$ and $10^2$.  With too low values of $\delta$, we observed numerical errors in the Cholesky decomposition required for constructing the \our dual.
\item Othewise: The ten $\delta$ values are logspaced between $10^{-2}$ and $10^2$, matching~\cite{immer2021improving}.
\end{itemize}

@@

\subsection{Supervised learning on image data sets}
\label{app:image}
% Each experiment with the image data sets consisted of the following steps.
% First, we trained a neural network (MLP for MNIST, CNN for FMNIST) using a prior precision $\delta$.
% After training, we performed inference by calculating the dual parameters.
% Finally, we evaluated the predictive performance of \our on the validation set.
% Similarly to the supervised UCI benchmark tasks, a single image experiment consists of the following steps: first, we train a neural network (a MLP for MNIST, a CNN for FMNIST) in the classification task, using a prior precision $\delta$. After training, we prepare the \our method for predictions by computing the dual parameters, and then evaluate the validation set predictive performance of \our. While in principle our method's predictions consist of the GP mean and variance, we include results for using the neural network prediction as a mean instead. For comparison, we gather the GP subset method results for both the GP and neural network means. This experiment set-up is then replicated for a range of prior precision values and learning rates, as we perform a grid search over these hyperparameters by optimizing for the validation set NLPD. We run the experiment on $5$ seeds and pick the hyperparameter by the best mean performance over the seeds. The results reported in \cref{tbl:imagesuper} used a learning rate of $10^{-4}$, and a prior precision of $\delta = 0.0013$ for FMNIST and $\delta = 0.0001$ for MNIST \todo{double check}.

@@

\subsection{Updating the network representation in continual learning}
\label{app:cl-exp}

In \cref{sec:cl-exp} in the main paper, we showed how the regularizer described in \cref{sec:sequential} can be used in continual learning to retain a compact representation of the neural network. We conducted our experiments on three common CL data sets \cite{de2021continual, pan2020continual, rudner2022continual}: 
\begin{itemize}
	\item Split-MNIST (S-MNIST), that consists in five binary classification tasks each on pair on MNIST classes (\ie, \digit{0}~and~\digit{1}, \digit{2}~and~\digit{3}, \dots, \digit{8}~and~\digit{9});
	\item Split-FashionMNIST (S-FMNIST), composed of five binary tasks as S-MNIST, but using instead Fashion-MNIST;
	\item Permuted-MNIST (P-MNIST), that consists of ten tasks, where each one is a ten-way classification task on the entire MNIST data set, where the images' pixels are randomly permuted.
\end{itemize}

Following \cite{pan2020continual, rudner2022continual}, we used a two-layer MLP with 256 hidden units and ReLU activations for S-MNIST and S-FMNIST and a two-layer MLP with 100 hidden units and ReLU activations for P-MNIST, to ensure consistency with the other methods. In our experiments we only consider the single-head (SH) setting. This means that for S-MNIST and S-FMNIST, we are dealing with a \emph{Class-Incremental Learning} (CIL) scenario, where for each task a pair of dijoint classes are presented to the model. The P-MNIST experiments falls instead into the \emph{Domain-Incremental Learning} (DIL) setting, where the model deals with a data-distribution shift, by progressively observing tasks with different permutations of the image pixels that need to be classified into ten different classes.

In our experiments, we compared the performance of the \our-based regularizer against two families of methods: weight-space regularization methods and function-space regularization ones.
For the weight-space methods, we considered Online-EWC~\citep{schwarz2018progress}, SI~\citep{zenke2017continual}, and VCL~\citep{nguyen2018variational}.  % In \cref{tbl:cl_table_1}, we considered VCL with random coresets, which uses coreset points t 
For the function-space methods, we considered DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual}.

For our method, DER, Online-EWC, and SI, we relied on the Mammoth framework proposed by \citep{buzzega2020dark}. On the other hand, we used the available open-source FROMP codebase to conduct the experiments with FROMP, and the S-FSVI codebase to run the experiments on S-FSVI and VCL. We selected the hyperparameters to use with all these methods trying to adapt the ones pointed out in the their original papers and the reference implementations. For this reason, we relied on SGD for DER, Online-EWC, and SI, which requires higher learning rates, compared to the others methods for which we used Adam, \ie, our \our regularizer, FROMP, S-FSVI, and VCL.

For the methods relying on a stored subset of training data for each task, we make sure that the number of data points stored are the same for each method. For DER, which does not select the rehearsal training samples at task boundaries, we cannot apply directly the definition of points per task ($M$), and thus we set the dimension of the reservoir buffer to be $T \cdot M$, \ie, total number of tasks times the points per task.

The results reported in \cref{tbl:cl_table_1}, are obtained using the following hyperparameters. The results are in-line with what has been reported in previous work, notably even better for some of the baseline methods. Thus, we consider the presented results to give a realistic picture of the relative performance of the different methods, and a demonstrating a clear case for applicability of the \our approach in continual learning.

%% Hyperparms used in the different experiments:
\paragraph{S-MNIST (40 pts./task):}
%\begin{itemize}
	\textbf{Online-EWC}: SGD, learning rate $0.03$, batch size  10, number of epochs 1, $\lambda=90$, $\gamma=1.0$.
	\textbf{SI}: SGD, learning rate 0.1, batch size 10, number of epochs 1, $c=1.0$, $\xi=0.9$.
	\textbf{VCL (with coresets)}: Adam, learning rate 0.001, batch size 256, number of epochs 100, selection method for coresets random choice.
	\textbf{DER}: SGD with learning rate 0.03, batch size 10, buffer batch size 10, number epochs 1, $\alpha=0.3$, reservoir buffer size 200.
	\textbf{FROMP}: Adam, learning rate $1 \cdot 10^{-4}$, batch size 32, number of epochs 10, $\tau = 10$, random selection for memorable points.
	\textbf{S-FSVI}: Adam, learning rate $5 \cdot 10^{-4}$, batch size 128, number of epochs 10.
	\textbf{SFR}: Adam, learning rate $3 \cdot 10^{-4}$, batch size 32, $\tau=1$, $\delta=0.0001$, number of epochs 1.
%\end{itemize}

\paragraph{S-MNIST (200 pts./task):}
%\begin{itemize}
	\textbf{Online-EWC}: SGD, learning rate $0.03$, batch size  10, number of epochs 1, $\lambda=90$, $\gamma=1.0$.
	\textbf{SI}: SGD, learning rate 0.1, batch size 10, number of epochs 1, $c=1.0$, $\xi=0.9$.
	\textbf{VCL (with coresets)}: Adam, learning rate 0.001, batch size 256, number of epochs 100, selection method for coresets random choice.
	\textbf{DER}: SGD, learning rate 0.03, batch size 10, buffer batch size 10, number epochs 1, $\alpha$ 0.2, reservoir buffer size 1000.
	\textbf{FROMP}: Adam, learning rate $1 \cdot 10^{-4}$, batch size 32, number of epochs 10, $\tau = 10$, random selection for memorable points.
	\textbf{S-FSVI}: Adam, learning rate $5 \cdot 10^{-4}$, batch size 128, number of epochs 80.
	\textbf{SFR}: Adam, learning rate $1 \cdot 10^{-4}$, batch size 32, $\tau=0.5$, $\delta=1 \cdot 10^{-5}$, number of epochs 5.
%\end{itemize}

%% Placeholders for now
\paragraph{S-FMNIST (200 pts./task):}
%\begin{itemize}
	\textbf{Online-EWC}: SGD, learning rate $0.03$, batch size  10, number of epochs 5, $\lambda=90$, $\gamma=1.0$.
	\textbf{SI}: SGD, learning rate 0.1, batch size 10, number of epochs 5, $c=1.0$, $\xi=0.9$.
	\textbf{VCL (with coresets)}: Adam, learning rate $1 \cdot 10^{-3}$, batch size 256, number of epochs 100, selection method for coresets random choice.
	\textbf{DER}: SGD, learning rate 0.03, batch size 10, buffer batch size 10, number epochs 5, $\alpha =  0.2$, reservoir buffer size 200.
	\textbf{FROMP}: Adam, learning rate $1 \cdot 10^{-4}$, batch size 32, number of epochs 10, $\tau = 10$, random selection for memorable points.
	\textbf{S-FSVI}: Adam, learning rate $5 \cdot 10^{-4}$, batch size 128, number of epochs 60.
	\textbf{SFR}: Adam, learning rate $3 \cdot 10^{-4}$, batch size 32, $\tau=1.0$, $\delta=0.0001$, number of epochs 5.
%\end{itemize}

\paragraph{P-MNIST (200 pts./task):}
%\begin{itemize}
	\textbf{Online-EWC}: SGD, learning rate $0.1$, batch size 128, number of epochs 10, $\lambda=0.7$, $\gamma=1.0$.
	\textbf{SI}: SGD, learning rate 0.1, batch size 138, number of epochs 10, $c=0.5$, $\xi=1.0$. 
	\textbf{VCL (with coresets)}: Adam, learning rate 0.001, batch size 256, number of epochs 100, selection method for coresets random choice.
	\textbf{DER}:  SGD, learning rate 0.01, batch size 128, buffer batch size 128, number epochs 10, $\alpha=0.3$, reservoir buffer size 2000.
	\textbf{FROMP}: Adam, learning rate $1 \cdot 10^{-3}$, batch size 128, number of epochs 10, $\tau = 0.5$, lambda descend selection for memorable points.
	\textbf{S-FSVI}: Adam, learning rate $5 \cdot 10^{-4}$, batch size 128, number of epochs 10.
	\textbf{SFR}: Adam, learning rate $3 \cdot 10^{-4}$, batch size 64, $\tau=1.0$, $\delta=0.0001$. 
%\end{itemize}

Additionally, for the ablation experiments for \our, we used the same hyperparameters as reported for the full approach.
